# NFL 

## WBs
```{r}
library(readr)
library(randomForest)
library(caret)
library(tidyr)
library(pdp)
library(dplyr)

```


```{r}
# Load the offensive yearly data
data <- read_csv("offense_yearly_data.csv")

# Sort the dataset by player name and season
data <- data[order(data$name, data$season), ]
```


```{r}
names(data)[names(data) == "height_cm"] <- "height_in"
data <- data[, !names(data) %in% c("height_ft")]
```


```{r}
# Split the dataset by position
data_split <- split(data, data$position)

# Access the WR data
wr_data <- data_split$WR
```




```{r}
# Identify zero-variance columns
zero_var_cols <- nearZeroVar(wr_data)

# Check and remove zero-variance columns, excluding 'position'
zero_var_colnames <- colnames(wr_data)[zero_var_cols]
cols_to_keep <- names(wr_data)[!(names(wr_data) %in% zero_var_colnames) | names(wr_data) == "position"]
wr_data <- wr_data[, cols_to_keep]

```

```{r}
# Filter the data for the 2023 season to save it before we kill it
wr_2023 <- subset(wr_data, season == 2023)

```

```{r}
# Create new columns for the next season's fantasy points
wr_data$next_fantasy_points <- ave(wr_data$fantasy_points, wr_data$name, FUN = function(x) c(x[-1], NA))
wr_data$next_fantasy_points_ppr <- ave(wr_data$fantasy_points_ppr, wr_data$name, FUN = function(x) c(x[-1], NA))

# View the first few rows of the updated data
head(data)

```

```{r}
wr_data <- wr_data[wr_data$rookie_season != 0, ]

```


```{r}
# Count the number of NA values in each column
na_table <- colSums(is.na(wr_data))
print(na_table)

# For now, remove all rows with NAs
wr_data <- na.omit(wr_data)

# Drop the next_fantasy_points_ppr column
wr_data <- wr_data[, !names(wr_data) %in% "next_fantasy_points_ppr"]

```


```{r}
# Build Random Forest model
wrrf_model <- randomForest(next_fantasy_points ~ ., data = wr_data, ntree = 500, importance = TRUE)
```


```{r}
# Create a simplified dataset for linear regression
lm_data <- wr_data[, !names(wr_data) %in% c("name", "id", "position")]


# Initialize a vector to store RMSPE values for each run
rmspe <- c()

for(i in 1:10) {
  
  # Resample the dataset with replacement
  ind <- unique(sample(nrow(lm_data), nrow(lm_data), replace = TRUE))
  train <- lm_data[ind, ]
  test <- lm_data[-ind, ]
  
  # Build Linear Regression model
  lm_model <- lm(next_fantasy_points ~ ., data = train)
  
  # Predict on the test data
  yhat <- predict(lm_model, newdata = test)
  
  # Calculate RMSPE
  rmspe[i] <- sqrt(mean((test$next_fantasy_points - yhat)^2))
}

# Calculate the mean RMSPE over the 10 runs
mean_rmspe <- mean(rmspe)
print(paste("Mean RMSPE over 10 runs:", mean_rmspe))

# Optionally, print the RMSPE for each run to see variability
print(rmspe)

```


```{r}
# Evaluate feature importance
importance_values <- importance(wrrf_model)

# Sort and extract the top 10 features by %IncMSE and IncNodePurity
top10_mse <- importance_values[order(importance_values[, "%IncMSE"], decreasing = TRUE), ][1:10, ]
top10_purity <- importance_values[order(importance_values[, "IncNodePurity"], decreasing = TRUE), ][1:10, ]

# Visualize top 10 features by %IncMSE
par(mar = c(5, 12, 4, 2))  # Increase margins for labels
barplot(top10_mse[, "%IncMSE"], names.arg = rownames(top10_mse),
        main = "Top 10 Features by %IncMSE (Mean Decrease in Accuracy)",
        las = 2, col = "blue", horiz = TRUE, cex.names = 0.8)

# Visualize top 10 features by IncNodePurity
barplot(top10_purity[, "IncNodePurity"], names.arg = rownames(top10_purity),
        main = "Top 10 Features by IncNodePurity (Mean Decrease in Gini)",
        las = 2, col = "red", horiz = TRUE, cex.names = 0.8)

```

```{r}
oob_mse <- wrrf_model$mse[wrrf_model$ntree]
oobrmse <- sqrt(oob_mse)
# Print the OOB MSE
print(paste("OOB rMSE:", oobrmse))
```
```{r}
# Extract OOB predictions
oob_predictions <- wrrf_model$predicted

# Calculate the absolute errors
absolute_errors <- abs(oob_predictions - wrrf_model$y)

# Calculate the Mean Absolute Error (MAE)
oob_mae <- mean(absolute_errors)

# Print the OOB MAE
print(paste("OOB MAE:", oob_mae))

```



```{r}

# Predict 2024 fantasy points for 2023 WRs
predicted_2024_fantasy_points <- predict(wrrf_model, newdata = wr_2023)

# Create a data frame with the names and the predicted 2024 fantasy points
results_wr <- data.frame(name = wr_2023$name, predicted_2024_fantasy_points = predicted_2024_fantasy_points)
print(results_wr)

```


mse of specific seasons?
we need to factor in the qbs and teamates of the players

## RBs

```{r}
# Access the RB data
rb_data <- data_split$RB
```

```{r}
# Identify zero-variance columns
zero_var_cols <- nearZeroVar(rb_data)

# Check and remove zero-variance columns, excluding 'position'
zero_var_colnames <- colnames(rb_data)[zero_var_cols]
cols_to_keep <- names(rb_data)[!(names(rb_data) %in% zero_var_colnames) | names(rb_data) == "position"]
rb_data <- rb_data[, cols_to_keep]

rb_2023 <- subset(rb_data, season == 2023)

```

```{r}
# Create new columns for the next season's fantasy points
rb_data$next_fantasy_points <- ave(rb_data$fantasy_points, rb_data$name, FUN = function(x) c(x[-1], NA))
rb_data$next_fantasy_points_ppr <- ave(rb_data$fantasy_points_ppr, rb_data$name, FUN = function(x) c(x[-1], NA))

# View the first few rows of the updated data
head(rb_data)

```

```{r}
rb_data <- rb_data[rb_data$rookie_season != 0, ]

```


```{r}
# Count the number of NA values in each column
na_table <- colSums(is.na(rb_data))
print(na_table)

# For now, remove all rows with NAs
rb_data <- na.omit(rb_data)

# Drop the next_fantasy_points_ppr column
rb_data <- rb_data[, !names(rb_data) %in% "next_fantasy_points_ppr"]

```

```{r}
# Build Random Forest model
rbrf_model <- randomForest(next_fantasy_points ~ ., data = rb_data, ntree = 500, importance = TRUE)

```

```{r}
# Create a simplified dataset for linear regression
lm_data <- rb_data[, !names(rb_data) %in% c("name", "id", "position")]

# Initialize a vector to store RMSPE values for each run
rmspe <- c()

for(i in 1:10) {
  
  # Resample the dataset with replacement
  ind <- unique(sample(nrow(lm_data), nrow(lm_data), replace = TRUE))
  train <- lm_data[ind, ]
  test <- lm_data[-ind, ]
  
  # Build Linear Regression model
  lm_model <- lm(next_fantasy_points ~ ., data = train)
  
  # Predict on the test data
  yhat <- predict(lm_model, newdata = test)
  
  # Calculate RMSPE
  rmspe[i] <- sqrt(mean((test$next_fantasy_points - yhat)^2))
}

# Calculate the mean RMSPE over the 10 runs
mean_rmspe <- mean(rmspe)
print(paste("Mean RMSPE over 10 runs:", mean_rmspe))

# Optionally, print the RMSPE for each run to see variability
print(rmspe)

```

```{r}
# Evaluate feature importance
importance_values <- importance(rbrf_model)

# Sort and extract the top 10 features by %IncMSE and IncNodePurity
top10_mse <- importance_values[order(importance_values[, "%IncMSE"], decreasing = TRUE), ][1:10, ]
top10_purity <- importance_values[order(importance_values[, "IncNodePurity"], decreasing = TRUE), ][1:10, ]

# Visualize top 10 features by %IncMSE
par(mar = c(5, 12, 4, 2))  # Increase margins for labels
barplot(top10_mse[, "%IncMSE"], names.arg = rownames(top10_mse),
        main = "Top 10 Features by %IncMSE (Mean Decrease in Accuracy)",
        las = 2, col = "blue", horiz = TRUE, cex.names = 0.8)

# Visualize top 10 features by IncNodePurity
barplot(top10_purity[, "IncNodePurity"], names.arg = rownames(top10_purity),
        main = "Top 10 Features by IncNodePurity (Mean Decrease in Gini)",
        las = 2, col = "red", horiz = TRUE, cex.names = 0.8)

```

```{r}
# Calculate OOB MAE
# Extract OOB predictions
oob_predictions <- rbrf_model$predicted

# Calculate the absolute errors
absolute_errors <- abs(oob_predictions - rbrf_model$y)

# Calculate the Mean Absolute Error (MAE)
oob_mae <- mean(absolute_errors)
print(paste("OOB MAE:", oob_mae))

```
```{r}
# Predict 2024 fantasy points for 2023 RBs
predicted_2024_fantasy_points <- predict(rbrf_model, newdata = rb_2023)

# Create a data frame with the names and the predicted 2024 fantasy points
results_rb <- data.frame(name = rb_2023$name, predicted_2024_fantasy_points = predicted_2024_fantasy_points)
print(results_rb)

```

## QBs 

```{r}
# Access the QB data
qb_data <- data_split$QB
```

```{r}

# Identify zero-variance columns
zero_var_cols <- nearZeroVar(qb_data)

# Check and remove zero-variance columns
zero_var_colnames <- colnames(qb_data)[zero_var_cols]
qb_data <- qb_data[, !names(qb_data) %in% zero_var_colnames]


# Filter the data for the 2023 season
qb_2023 <- subset(qb_data, season == 2023)
# Create new columns for the next season's fantasy points
qb_data$next_fantasy_points <- ave(qb_data$fantasy_points, qb_data$name, FUN = function(x) c(x[-1], NA))

# Remove rows where rookie_season is 0
qb_data <- qb_data[qb_data$rookie_season != 0, ]

# Remove rows with NAs
qb_data <- na.omit(qb_data)

```

```{r}
# Build Random Forest model
qbrf_model <- randomForest(next_fantasy_points ~ ., data = qb_data, ntree = 500, importance = TRUE)

```

```{r}
# Evaluate feature importance
importance_values <- importance(qbrf_model)

# Sort and extract the top 10 features by %IncMSE and IncNodePurity
top10_mse <- importance_values[order(importance_values[, "%IncMSE"], decreasing = TRUE), ][1:10, ]
top10_purity <- importance_values[order(importance_values[, "IncNodePurity"], decreasing = TRUE), ][1:10, ]

# Visualize top 10 features by %IncMSE
par(mar = c(5, 12, 4, 2))  # Increase margins for labels
barplot(top10_mse[, "%IncMSE"], names.arg = rownames(top10_mse),
        main = "Top 10 Features by %IncMSE (Mean Decrease in Accuracy)",
        las = 2, col = "blue", horiz = TRUE, cex.names = 0.8)

# Visualize top 10 features by IncNodePurity
barplot(top10_purity[, "IncNodePurity"], names.arg = rownames(top10_purity),
        main = "Top 10 Features by IncNodePurity (Mean Decrease in Gini)",
        las = 2, col = "red", horiz = TRUE, cex.names = 0.8)

```

```{r}
# Calculate OOB rmse
oob_rmse <- sqrt(qbrf_model$mse[qbrf_model$ntree])
# Extract OOB predictions
oob_predictions <- qbrf_model$predicted

# Calculate the absolute errors
absolute_errors <- abs(oob_predictions - qbrf_model$y)

# Calculate the Mean Absolute Error (MAE)
oob_mae <- mean(absolute_errors)
print(paste("OOB MAE:", oob_mae))
oob_rmse
```


```{r}
# Predict 2024 fantasy points for 2023 QBs
predicted_2024_fantasy_points <- predict(qbrf_model, newdata = qb_2023)

# Create a data frame with the names and the predicted 2024 fantasy points
results_qb <- data.frame(name = qb_2023$name, predicted_2024_fantasy_points = predicted_2024_fantasy_points)
print(head(results_qb))

```










## Plots and Analysis

```{r}
# For QB model (qbrf_model)
pdp_qb_games <- partial(qbrf_model, pred.var = "games", plot = TRUE, 
                        main = "PDP for Games (QB Random Forest Model)",
                        xlab = "Games", ylab = "Partial Dependence")



# For WR model (wrrf_model)
pdp_wr_games <- partial(wrrf_model, pred.var = "games", plot = TRUE, 
                        main = "PDP for Games (WR Random Forest Model)",
                        xlab = "Games", ylab = "Partial Dependence")

# For RB model (rbrf_model)
pdp_rb_games <- partial(rbrf_model, pred.var = "games", plot = TRUE, 
                        main = "PDP for Games (RB Random Forest Model)",
                        xlab = "Games", ylab = "Partial Dependence")

pdp_qb_games
pdp_wr_games
pdp_rb_games
```

Games played is being tracked by other statistics... i dont give a crap about games played i only care about your stats at the end of the season...

## Lists to do

need to look at if multiple years back is good idea(i dont think so because then can no use 2nd yr players), and to be able to add teamates. also look if we are good for the seasons that there are 17 games. break the seasons down into quarters. 

## team mates 

```{r}
library(nflfastR)
library(dplyr)

```


```{r}
# Fetch rosters for 2024
rosters <- fast_scraper_roster(season = 2024)

# Filter WR and RB positions
wr_rosters <- rosters %>% filter(position == "WR")
rb_rosters <- rosters %>% filter(position == "RB")

# Rename columns to match merging requirements
colnames(wr_2023)[colnames(wr_2023) == "team.x"] <- "team"
colnames(rb_2023)[colnames(rb_2023) == "team.x"] <- "team"

# Merge WR data with rosters
wr_teams <- merge(wr_2023, wr_rosters, by.x = "name", by.y = "full_name", all.x = TRUE)
colnames(wr_teams)[colnames(wr_teams) == "team.x"] <- "team" # Use team.x for original team
colnames(wr_teams)[colnames(wr_teams) == "team.y"] <- "new_team" # Use team.y for updated team
wr_teams <- wr_teams[, c("name", "team")]

# Merge RB data with rosters
rb_teams <- merge(rb_2023, rb_rosters, by.x = "name", by.y = "full_name", all.x = TRUE)
colnames(rb_teams)[colnames(rb_teams) == "team.x"] <- "team" # Use team.x for original team
colnames(rb_teams)[colnames(rb_teams) == "team.y"] <- "new_team" # Use team.y for updated team
rb_teams <- rb_teams[, c("name", "team")]


```

create wr_2023-2024 that has the 2023 stats that have the players on there correct teams. we will do the same for the rb data
```{r}
# Load the necessary libraries
library(readr)
library(randomForest)
library(caret)
library(tidyr)
library(pdp)
library(dplyr)
library(nflfastR)

# Load the offensive yearly data
data <- read_csv("offense_yearly_data.csv")

# Sort the dataset by player name and season
data <- data[order(data$name, data$season), ]

# Adjust columns
names(data)[names(data) == "height_cm"] <- "height_in"
data <- data[, !names(data) %in% c("height_ft")]

# Split the dataset by position
data_split <- split(data, data$position)

# Access the WR and RB data
wr_data <- data_split$WR
rb_data <- data_split$RB

# Identify zero-variance columns and remove them, excluding 'position'
# For WR data
zero_var_cols_wr <- nearZeroVar(wr_data)
zero_var_colnames_wr <- colnames(wr_data)[zero_var_cols_wr]
cols_to_remove_wr <- zero_var_colnames_wr[zero_var_colnames_wr != "position"]
wr_data <- wr_data[, !names(wr_data) %in% cols_to_remove_wr]

# For RB data
zero_var_cols_rb <- nearZeroVar(rb_data)
zero_var_colnames_rb <- colnames(rb_data)[zero_var_cols_rb]
cols_to_remove_rb <- zero_var_colnames_rb[zero_var_colnames_rb != "position"]
rb_data <- rb_data[, !names(rb_data) %in% cols_to_remove_rb]

# Assuming the previous steps up to filtering the 2023 data are done
# Filter the data for the 2023 season
wr_2023 <- subset(wr_data, season == 2023)
rb_2023 <- subset(rb_data, season == 2023)

# Fetch current rosters for 2024
rosters <- fast_scraper_roster(season = 2024)
wr_rosters <- rosters %>% filter(position == "WR")
rb_rosters <- rosters %>% filter(position == "RB")

# For WR data
wr_rosters_temp <- wr_rosters[, c("full_name", "team")]
names(wr_rosters_temp)[names(wr_rosters_temp) == "team"] <- "team_roster"
wr_2023_2024 <- merge(wr_2023, wr_rosters_temp, by.x = "name", by.y = "full_name", all.x = TRUE)
# Replace NA in team with team_roster if available
wr_2023_2024$team <- ifelse(is.na(wr_2023_2024$team_roster), 
                           wr_2023_2024$team, 
                           wr_2023_2024$team_roster)
wr_2023_2024 <- wr_2023_2024[, !names(wr_2023_2024) %in% "team_roster"]

# For RB data
rb_rosters_temp <- rb_rosters[, c("full_name", "team")]
names(rb_rosters_temp)[names(rb_rosters_temp) == "team"] <- "team_roster"
rb_2023_2024 <- merge(rb_2023, rb_rosters_temp, by.x = "name", by.y = "full_name", all.x = TRUE)
# Replace NA in team with team_roster if available
rb_2023_2024$team <- ifelse(is.na(rb_2023_2024$team_roster), 
                           rb_2023_2024$team, 
                           rb_2023_2024$team_roster)
rb_2023_2024 <- rb_2023_2024[, !names(rb_2023_2024) %in% "team_roster"]

# Now, wr_2023_2024 and rb_2023_2024 should have the correct teams for the players in 2023, with fallback to original team if NA.


```
now in wr_data and rb we need to look to include teamate stats. (we will group them for the 2023/2024 one later) we will create collums that have a the teamates states. to determine a teamate use the team and the season... now we obviously cant use every teamate. so for now we wil use the best qb on that team the 2 best wrs(that isnt themselves) and the best rb... now we have to determine what best is. best for qb will be pass yards. best for rbs will be rush yards. and best for wrs will be receving yards we may need to restructure it from the begining a bit because some colloms where dropped but u figure that out


```{r}
# Load the necessary libraries
library(readr)
library(dplyr)
library(tidyr)

# Load the offensive yearly data
data <- read_csv("offense_yearly_data.csv")

# Sort the dataset by player name and season
data <- data[order(data$name, data$season), ]

# Adjust columns
names(data)[names(data) == "height_cm"] <- "height_in"  # Rename column
data <- data[, !names(data) %in% "height_ft"]          # Remove height_ft column

# Split the dataset by position
data_split <- split(data, data$position)

# Access the WR, RB, and QB data
wr_data <- data_split$WR
rb_data <- data_split$RB
qb_data <- data_split$QB



# Load the necessary libraries
library(readr)
library(dplyr)

# Load the offensive yearly data
data <- read_csv("offense_yearly_data.csv")

# Group by season and team to get a column with the names of all team members
get_team_members <- function(data) {
  data %>%
    group_by(season, team) %>%
    summarise(team_members = paste(name, collapse = ", ")) %>%
    ungroup()
}

# Apply the function to WR, RB, and QB data
wr_team_members <- get_team_members(data_split$WR)
rb_team_members <- get_team_members(data_split$RB)
qb_team_members <- get_team_members(data_split$QB)



```

```{r}
# Function to get team members
get_team_members <- function(data) {
  data %>%
    group_by(season, team) %>%
    summarise(team_members = paste(name, collapse = ", ")) %>%
    ungroup()
}

# Apply the function to existing data
wr_team_members <- get_team_members(data_split$WR)
rb_team_members <- get_team_members(data_split$RB)
qb_team_members <- get_team_members(data_split$QB)

# Function to get top WRs by team
get_top_wr_team_members <- function(ranked_wr_data) {
  # Sort data
  ranked_wr_data <- ranked_wr_data[order(ranked_wr_data$season, 
                                         ranked_wr_data$team, 
                                         ranked_wr_data$rank), ]
  
  # Function to get nth value safely
  get_nth <- function(x, n) {
    if(length(x) >= n) return(x[n])
    return(NA)
  }
  
  # Split data by season and team
  split_data <- split(ranked_wr_data, 
                      list(ranked_wr_data$season, ranked_wr_data$team))
  
  # Create result data frame
  result_list <- lapply(split_data, function(group) {
    # If the group is empty, return NULL
    if (nrow(group) == 0) {
      return(NULL) 
    }
    
    data.frame(
      season = unique(group$season),
      team = unique(group$team),
      WR1_name = get_nth(group$name, 1),
      WR1_yards = get_nth(group$receiving_yards, 1),
      WR2_name = get_nth(group$name, 2),
      WR2_yards = get_nth(group$receiving_yards, 2),
      WR3_name = get_nth(group$name, 3),
      WR3_yards = get_nth(group$receiving_yards, 3),
      WR4_name = get_nth(group$name, 4),
      WR4_yards = get_nth(group$receiving_yards, 4),
      WR5_name = get_nth(group$name, 5),
      WR5_yards = get_nth(group$receiving_yards, 5),
      stringsAsFactors = FALSE
    )
  })
  
  # Filter out NULL entries
  result_list <- Filter(Negate(is.null), result_list)
  
  # Combine all data frames into one
  result <- do.call(rbind, result_list)
  
  # Create team_key
  result$team_key <- paste(result$team, result$season, sep = "_")
  
  # Reorder columns
  col_order <- c("team_key", "season", "team", 
                 names(result)[!names(result) %in% c("team_key", "season", "team")])
  result <- result[, col_order]
  
  rownames(result) <- NULL
  return(result)
}

# Create a function to calculate dense rank
dense_rank_desc <- function(x) {
  # Sort unique values in descending order
  sorted_unique <- sort(unique(x), decreasing = TRUE)
  # Create rank mapping
  rank_map <- seq_along(sorted_unique)
  names(rank_map) <- sorted_unique
  # Return ranks
  rank_map[as.character(x)]
}

# Split data by season and team
splits <- split(wr_data, list(wr_data$season, wr_data$team))

# Function to process each group
process_group <- function(group) {
  if(nrow(group) == 0) return(NULL)
  
  # Calculate rank within group
  group$rank <- dense_rank_desc(group$receiving_yards)
  
  # Select only needed columns
  cols_to_keep <- c("season", "team", "name", "receiving_yards", "rank")
  group <- group[, cols_to_keep[cols_to_keep %in% names(group)]]
  
  return(group)
}

# Apply processing to each group and combine results
ranked_wr_data <- do.call(rbind, lapply(splits, process_group))

# Reset row names
rownames(ranked_wr_data) <- NULL

# Create WR summary
wr_summary <- get_top_wr_team_members(ranked_wr_data)

# Check structure
str(wr_summary)
```



## Testing


```{r}
# Create team_key and merge with wr_summary
wr_data_final <- wr_data
wr_data_final$team_key <- paste(wr_data_final$team, wr_data_final$season, sep = "_")
wr_data_final <- merge(wr_data_final, wr_summary, by = "team_key", all.x = TRUE)

# Rename columns and create initial wr_dataf
wr_dataf <- wr_data_final
names(wr_dataf)[names(wr_dataf) == "season.x"] <- "season"
names(wr_dataf)[names(wr_dataf) == "team.x"] <- "team"
names(wr_dataf)[names(wr_dataf) == "name"] <- "player_name"

# Remove ID column if it exists
if("id" %in% names(wr_dataf)) {
    wr_dataf <- wr_dataf[, !names(wr_dataf) == "id"]
}

# Subset 2023 data
wr_2023 <- wr_dataf[wr_dataf$season == 2023, ]

# Create next season's fantasy points
create_next_points <- function(points, player) {
    ave(points, player, FUN = function(x) c(x[-1], NA))
}
wr_dataf$next_fantasy_points <- create_next_points(wr_dataf$fantasy_points, wr_dataf$player_name)
wr_dataf$next_fantasy_points_ppr <- create_next_points(wr_dataf$fantasy_points_ppr, wr_dataf$player_name)

# Filter rookie seasons
wr_dataf <- wr_dataf[wr_dataf$rookie_season != 0, ]
wr_dataf <- wr_dataf[wr_dataf$rookie_season != wr_dataf$season, ]

# Remove next_fantasy_points_ppr
wr_dataf$next_fantasy_points_ppr <- NULL

# Remove zero-variance columns
n_distinct <- function(x) length(unique(x))
zero_var_cols <- names(wr_dataf)[sapply(wr_dataf, n_distinct) == 1]
wr_dataf <- wr_dataf[, !names(wr_dataf) %in% zero_var_cols]

# Remove non-predictive columns
columns_to_remove <- c("player_name", "team_key", "WR2_name", "WR1_name", 
                      "RB3_name", "RB2_name", "RB1_name", "QB_name", 
                      "season", "team", "season.y", "team.y")
wr_dataf <- wr_dataf[, !names(wr_dataf) %in% columns_to_remove]

# Train Random Forest model
rf_model <- randomForest(next_fantasy_points ~ ., 
                        ntree = 1200, 
                        data = wr_dataf, 
                        importance = TRUE, 
                        na.action = na.omit)

# Calculate performance metrics
oob_predictions <- rf_model$predicted
absolute_errors <- abs(oob_predictions - rf_model$y)
oob_mae <- mean(absolute_errors)
oob_mse <- rf_model$mse[1200]
oobrmse <- sqrt(oob_mse)

# Print metrics
print(paste("OOB MAE:", oob_mae))
print(paste("OOB rMSE:", oobrmse))

# Feature importance
importance_values <- importance(rf_model)
ordered_importance <- order(importance_values[, "%IncMSE"], decreasing = TRUE)
top10_mse <- importance_values[ordered_importance[1:10], ]

# Plot importance
par(mar = c(5, 12, 4, 2))
barplot(top10_mse[, "%IncMSE"], 
        names.arg = rownames(top10_mse),
        main = "Top 10 Features by %IncMSE (Enhanced Model)",
        las = 2, 
        col = "blue", 
        horiz = TRUE, 
        cex.names = 0.8)

# Save player names and clean 2023 data
nms <- wr_2023$player_name
wr_2023 <- wr_2023[, !names(wr_2023) %in% c(zero_var_cols, columns_to_remove)]

# Generate predictions
predicted_2024_fantasy_points <- predict(rf_model, newdata = wr_2023)

# Create and sort results
results_wr <- data.frame(
    name = nms,
    predicted_2024_fantasy_points = predicted_2024_fantasy_points
)
results_wr <- results_wr[order(results_wr$predicted_2024_fantasy_points, decreasing = TRUE), ]

# Print predictions
print("2024 WR Predictions (Enhanced Model):")
print(results_wr)
```









